\Large Generation of Mathematical Programming Representation for Discrete Event Simulation Models of Timed Petri Nets

### 2023-01-09 Answer to reviewers.tex

\usepackage[natbibapa,nodoi]{apacite}
\setlength\bibhang{12pt}
\renewcommand\bibliographytypesize{\fontsize{10}{12}\selectfont}

\usepackage{color}

\begin{document}
\noindent
Dear DE, AE and reviewers,

Thanks to your valuable comments, this submission has undergone a major revision. In this revised version, the MP models are developed based on general timed Petri nets (TPN) instead of the proposed representation of discrete event dynamic systems (DEVS) that was used in the previous submission. This change reduces the difficulty of applying the proposed approach and can be better understood by people familiar with DEVS. However, since the TPN that we consider does not model firing cancellation, this manuscript does not cover the systems with event cancellation, which the previous one did. The coauthors and I believe that the benefit of TPN-based MP formulation provides tighter connection to the DEVS literature and is more important than extending to system with cancellation.

@@

The computational complexity of using MP to optimize DEVS has been %noticed 
addressed in many works, and they dealt the issue with classic MP techniques such as Benders decomposition, linear approximation, columns-row generation. We cannot deny that the computation issue really exists, but the ways to overcome are also vast. Furthermore, the translation is bidirectional. One can translate simulation into MP to optimize a system, and also translate the resulting MP into simulation. The value of translating MP into simulation is that the research on MP (such as duality and sensitivity) can %dig
increase the value of simulation, which is currently usually regarded only as a performance evaluation tool.

In the following, we answer the comments in details.

\section{DE comments}
\textit{The merit of the proposed technique should be clarified either in a general setting or on specific classes of systems like Petri nets or finite automata, and comparisons with extant tools and algorithms should be made.}

\noindent
\textbf{Answer:} %In this submission, 
In the revised manuscript our approach %works
has been reformulated to have it working under general settings of timed Petri net. With an example of the $G/G/m$ queue, the differences between the proposed model and the state-of-the-art MP model are discussed.

@@

\section{AE Comments}
\textit{The contribution and usefulness of the proposed approach is not convincing and should be improved. }

\textit{1. The authors should convincingly demonstrate the benefit of the proposed MPR over other formalism like the automata and PN while also considering the complexity burden of solving the resulting mathematical program. }

\noindent
\textbf{Answer:} %In this submission, 
In the revised manuscript, the MPR formulation is equivalent to the simulation realization of TPN.

We justify the benefit and computational burden of the proposed approach as follows:

\textbf{The benefits of using MP models to represent discrete event systems are many. As already mentioned, %above,
when coupled with optimization, the MP-based algorithms can be faster and %gain 
reach better solution than black--box optimization algorithms. Furthermore, black--box approaches have limited capability in solving constrained optimization problems, %but 
while MP--based approaches can easily deal with them \citep{zhang2020models}. The vast theoretical and methodological results developed in the MP field can be introduced into the study of DEVS through simulation. For instance, using sensitivity analysis of MP models, gradient estimates can be easily derived as suggested in \citep{chan2008optimization}. In fact, %we do not proposed 
this paper does not propose to totally replace simulation with MP, but to enrich the toolbox for analyzing and optimizing DEVS. Finally, the application of MP models of DEVS is not limited to optimization but also to show formal properties of the system under study. For instance, \cite{basile2012k} proposed the sufficient and necessary condition of K-diagnosability of TPN based on MP analysis.}

\textbf{The major concern about the application of MP is the computational complexity, %and the solving procedure
as the solution procedure can be time--consuming or even unbearable. However, many approaches from optimization community are available to improve the efficiency. Linear programming approximation \citep{alfieri2012mathematical}, Benders decomposition \citep{weiss2015buffer}, row--column generation \citep{alfieri2020time}, have been studied to solve optimization problems in %reality
real contexts based on mathematical programming representation of DEVS.}
~\\

@@

\textit{2. Furthermore, the discussion about the application of the approach should be significantly strengthened. Currently, this is rather vague and there is no comparisons with existing approaches that are used to address these problems.}

\noindent
\textbf{Answer}: With an example of the $G/G/m$ queue, the differences between the proposed model and the state-of-the-art model are discussed.

@@

\section{Reviewer 1}
\textit{However, I think that the authors should clarify better their contribution with respect to the standard translation of other modelling formalisms into mathematical programming (see references below). From a certain point of view, mathematical programming is also a modelling formalism, even it is usually considered more a computational tool, anyway it is evident that it cannot replace finite state automata or Petri nets to model DES. Hence, my question is, there is a way to translate, as for example a PN using the proposed technique? If yes, is it more efficient? If not, what is the effective application domain of the proposed technique, in addition to queueing systems?}

\textit{References: \cite{basile2012k}, \cite{basile2012k}, \cite{bemporad1999control}}

\noindent
\textbf{Answer:} As the reviewer suggests, %this submission proposed 
in the revised manuscript we propose the translation of TPN into MP.

The difference between the proposed approach and the suggested references are %as follows.
detailed in the following.

[1] \cite{bemporad1999control} proposed a mixed integer programming modeling framework for hybrid systems, whose states are mixed integer, in discrete time, while this work deals with continuous--time discrete--state systems.

[2] An MP model was proposed to minimize the makespan of single-server manufacturing systems in \cite{di2020makespan}, %but this
while our work is more general and can model multi-server systems.

[3] \cite{basile2012k} proposed the model based on PN, not TPN, because %it deals with 
the paper is about diagnosability %other 
rather than time-dependent performance measures, such as waiting time or throughput.

The effective application domain of the proposed technique is basically related to optimization. The MP model of TPN %is easy to expand 
can be easily extended to an MP model of TPN optimization. Based on the proposed MP model, an example with the extended MP model to optimize the $G/G/m$ queue is presented in Section 4.3.

@@

\section{Reviewer 2}

\textit{1. The main criticisms I can to formulate about this MPR approach is that
(i) it produces only fixed-length simulations of a system, and (ii) it
results in high-dimension linear programs (O(NxE) where N is the
length of the execution and E the number of event types). This puts
stringent restrictions on the possible applications of this
encoding. For instance, steady-state behavior can only be
approximated, at the expense of very large MPR systems. This casts a
serious doubt about the usefulness of the whole approach, in
comparison with a DES simulator, implementing an operational
semantics of the particular DES considered in the paper.}

\textit{{For this reason, I recommend the rejection of the paper, and encourage the authors to replace part of the queueing network case-studies by an in-depth technical discussion of the possible applications of this MPR encoding.}}

\noindent
\noindent
\textbf{Answer:} The reviewer's comment is reasonable. We discuss the computational burden as follows in the submission:

\textbf{The major concern about the application of MP is the computational complexity, %and the solving procedure
as the solution procedure can be time--consuming or even unbearable. However, many approaches from optimization community are available to improve the efficiency. Linear programming approximation \citep{alfieri2012mathematical}, Benders decomposition \citep{weiss2015buffer}, row--column generation \citep{alfieri2020time}, have been studied to solve optimization problems in %reality
real contexts based on mathematical programming representation of DEVS.}

In the PhD thesis of one of the coauthors \citep{zhang2021resource}, a Benders decomposition (BD) algorithm solving integrated simulation-optimization mathematical programming models is proposed. In the BD procedure, the subproblem is the MP representation of simulation realization of TPN. The Benders cuts are derived from the dual solution of the subproblem. The thesis proposed an algorithm to solve the dual solution from simulation, so that the computational complexity is linear to the number of decision variables. The BD algorithm finds good feasible solutions fast and guarantees sample-path optimality in the long run. A working paper is draft to present the computationally efficient BD algorithm. The coauthors and I decide not to combine the algorithm in this manuscript, since we would like to make the topic concentrated on MPR modeling and easy to digest.

@@

~\\

\textit{2. Detailed Comments}

\textit{The paper reads clearly. There are only very few typos. Here is a list of typos: p 5, Section 2.3, line 2: ...proposed. This...}

\noindent
\textbf{Answer}: we have correct the typos.

@@

~\\

\textit{From a technical point of view, there is one thing that needs to be clarified, page 8: How do you choose constant M? Understandably, it should be large enough. However, I would expect to find in the paper a discussion about a lower bound on M.}

\noindent
\textbf{Answer}: In the subsections of Section 3, the value of $M$ is discussed each time it %appears
is used, as different values of $M$ are required in the constraints where $M$ is needed.

{Section 3.1: The value of $M$ in constraints (A1) and (A2) can be set to a value that is larger than the simulation time span as it is needed to bound the finishing time of transitions.}

{Section 3.2: Also in this case, the value of $M$ can be set to a value that is larger than the simulation time span for constraints (B1) and (B2) as 
it is used to bound the transition firing time. %The value of $M$ can be chosen as the upper bound of marking of place $p$ minus $(W^{p,t}-1)$ for constraints (B4). 
For constraints (B4), instead, the value of $M$ can be chosen as the upper bound of marking of place $p$ minus $(W^{p,t}-1)$.}

{Section 3.3: The value of $M$ in constraints (C3) and (C4) can be set to $K$, i.e., the total number of iterations. }

@@

~\\

\textit{The queueing networks case-studies are carefully detailed. Perhaps
some space could be gained here, and used Section 4, to allow a
detailed and technical discussion about the applications of the MPR
encoding.}

\noindent
\textbf{Answer:} We remove the case study on the queueing network and add new examples of comparison and the extended optimization formulation.

### 2023-01-09 main.tex
complete answer to reviewer

\textbf{THE PAPER FINISHES HERE?}
 %%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section{Introduction}
Discrete Event Simulation (DES) is one of the most used tools for performance evaluation of %discrete event dynamic
complex systems in manufacturing and service applications. Hence, simulation--optimization algorithms are widely used %developed for optimizing the parameters of such systems. 
when performance evaluation must be coupled with optimization, i.e., when the best system configuration, according to some criteria, has to be found meanwhile guaranteeing a given value of some performance measure.  
Most of the state--of--the--art simulation--optimization algorithms consider DES as a \textit{black--box} function, and the structure of DES models has been seldom studied. On the contrary, a minority of the simulation--optimization literature explores the structure of the DES models, and %that
such approaches are referred to as \textit{white--box} simulation--optimization. 
Under the black--box setting, simulation--optimization algorithms work in an iterative way, alternating simulation and optimization procedures, 
thus, possibly leading to computational inefficiency if the number of iterations and/or the computation time per iteration increases too much. 
%and always require %huge amount of 
%many iterations to explore the %search 
%feasible region of the optimization problem, %and can be
%thus possibly leading to %computationally inefficient. 
%computational inefficiency.
The benefit of white--box simulation--optimization is the saving of simulation budget or computation time since 
the optimization procedure is guided by the information contained in the structure of the DES model (\cite{zhang2020sample}). However, the barrier to the use of white--box simulation--optimization is modeling DES as white box, so that it eventually favors optimization.

%When there is already a DES model, 
The benefit of developing an MPR might be not obvious (especially when there is already a DES model) due to the extremely high complexity of solving it. 
%One may be confused about the motivation of this work, i.e., why one wants the MPR when he/she has already a simulation model at hand, since the complexity of solving mathematical programming is usually high. 
However, this work does not suggest solving the MPR directly with state-of-the-art optimizers. Instead, the optimal solution of the MPR can be obtained from a simulation run. The added value of the MPR is that the structure of the DES can be explored. With the vast theoretical and methodological results developed in the mathematical programming (MP) field, for instance sensitivity analysis as suggested by \cite{chan2008optimization}, the MPR of a simulation model favors the optimal design and control of the discrete event systems.

This work proposes a procedure to establish a white--box simulation model, which is an equivalent Mathematical Programming Representation (MPR) model, based on the well-known event--scheduling logic (\cite{law2014simulation}). %for  
The procedure is applicable to certain types of DES models, and the assumptions that the DES model should satisfy are also presented in this work.
%Specifically, the modeling procedure, the assumptions which the DES model has to satisfy so that it can be applied and some examples are discussed.

%Considering the literature on simulation--optimization, 
\cite{chan2008optimization} proposed a modeling framework to translate a DES model into an MPR model in a general sense. Their modeling framework is based on the Event Relationship Graph (ERG) of the system dynamics. To derive the MPR, % model, 
an ERG of the DES must be constructed and expanded to an elementary ERG (EERG) model, and a procedure can be applied to translate the EERG model into an MPR. %model. 
However, this procedure has some limitations. First, deriving an ERG is not an easy task, and the user must pay quite much attention to detect all the event relationships and complete the triggering conditions between each pair of related events. %That 
The difficulty of developing ERG limits the wide spread of this procedure. 
Second, the modeling procedure is case--by--case depending on the event relationships, which means that the user must analyze the event relationships one by one and identify which type of modeling, including the variables and constraints, he/she should apply for each event relationship. Consequently, an automatic procedure was not proposed. 
%which means the user has to first identify which situations he/she faces by analyzing the relationships between each couple of events in the EERG, and then choose the appropriate model, including the variables and constraints. 
This is quite %a burden
difficult, since EERG is an expansion of ERG; the resulting graph could be huge and writing down the complete MPR model could be even impossible.

This work proposes a procedure that does not need % without plotting 
the ERG and %the modeling procedure 
can be used to automatically generate the MPR in a general--purpose programming language. Despite %taking different paths
being different, the MPRs proposed in this work and \cite{chan2008optimization} lead to %the 
equivalent results, which, in turn, are both equivalent to a simulation realization. The procedure can also be applied to devise MPR of DES with event cancellation, which was not considered in the work of Chan and Schruben (2008).

Many works in the literature show the potentiality of this research direction. For instance, the gradient can be conveniently estimated from the simulation model, if the MPR is approximated into Linear Programming (LP) and %solving 
the dual can be conveniently obtained (\cite{chan2008optimization}). Moreover, if some of the parameters in the MPR are changed to decision variables, the MPR becomes an integrated simulation--optimization model. Solving the integrated model provides the optimal solution of the optimization problem (\cite{matta2008simulation}). MP--based algorithms, such as linear programming approximation (\cite{alfieri2012mathematical}), Benders decomposition (\cite{weiss2015buffer}), row-column generation (\cite{alfieri2020time}), have been applied to improve the efficiency of %solving the 
integrated MP model solution.

The application of MPR--based simulation--optimization approaches is usually found in operations management of manufacturing and service systems. %The integrated simulation--optimization model has proved itself to be well suited in solving the buffer allocation problem \citep{zhang2020BAP}. %The most efficient approach to finding the sample--path global optimal of the buffer allocation problem of serial production line is developed based on it \citep{zhang2020BAP}. 
%Thanks to the flexibility of DES in evaluating complex systems, the buffer allocation problem of production systems with complex blocking mechanism, such as kanban control, base stock control, extended kanban control, can be managed \citep{pedrielli2015integrated}. Thanks to the flexibility of MPR in modeling optimization problems, problems involving real--valued decision variables such as optimal production rate \citep{tan2015mathematical}, bottleneck detection and throughput improvement problem \citep{zhang2020models} have all been well addressed and the sample--path global optimal solution can be obtained. %It is worthy to notice that, 
The flexibility of DES for complex system evaluation and of MPR for modeling optimization problems, allowed the integrated simulation-optimization approach to be effectively applied to buffer allocation problems (\cite{zhang2020BAP}), even with complex blocking mechanisms (\cite{pedrielli2015integrated}) and to problems involving real value decision variables (\cite{tan2015mathematical,zhang2020models}). 
Before the above-mentioned works were proposed, there were many state--of--the--art heuristic approaches addressing those problems, but without any guarantee of global or local optimality. Thus, the development of MPR--based simulation--optimization has made its contribution in the research area of manufacturing and service system optimization.

The contributions of this work are several and refer to different aspects. First, it proposes an MPR of DES from event-scheduling execution logic, which is the foundation of many DES implementations. Thus, the procedure does not require much extra effort once one has an event-scheduling execution logic implemented, while the state-of-the-art approach requires the DES modeled with ERG. The proposed approach can be easily automated (the source code has been uploaded in a repository, as explained in the following).  Second, it proposes the MPR of event cancellation, which is usually used to model queue with abandonment or servers with failures but has never been studied in the literature. This work will be a foundation of developing math-programming-based simulation-optimization algorithm, based on the vast literature in mathematical programming filed, for instance, the gradient estimation. 
%Third, the vast literature in mathematical programming field can be applied to the resulting MPR, for instance, for gradient estimation. Last, 
The proposed MPR can also be easily transformed into an MPR for optimizing the design parameters of the DES, which are common optimization problems in operations management field.

The rest of the paper is organized as follows. Section \ref{sec:MPR} describes the generation of the MPR of a DES model, including a brief introduction of event-scheduling algorithm, the assumptions for applying the proposed procedure, a modified event-scheduling algorithm, and the MPR itself automatically generated based on the model. Section \ref{sec:app} shows several examples of DES and the generated MPR, whose equivalence has been validated. Result discussion and conclusion are reported in Section \ref{sec:discussion} and \ref{sec:conclusion}, respectively.

\section{MPR generation procedure} \label{sec:MPR}
\subsection{Event-scheduling execution logic of DES } \label{sec:SimAlgo}
The event-scheduling approach is the logic behind all the major DES software and used by practitioners when developing simulation codes with general purpose languages (\cite{law2014simulation}). For sake of completeness, the logic is briefly described. The fundamental elements are the \textit{system state} and the \textit{events}. The system state $\textbf{s}$ is a set of state variables $s$ needed to describe the system at a particular time. An event $\xi$ is composed by a collection of four objects, $(CS^{\xi}(\textbf{s}),\ CC^{\xi}(\textbf{s}),\ F^{\xi}(\textbf{s}),\ T^{\xi})$, where:
\begin{eqnarray}
	CS^{\xi}(\textbf{s}): && \text{condition to schedule}\nonumber\\
	CC^{\xi}(\textbf{s}): && \text{condition to cancel}\nonumber\\
	F^{\xi}(\textbf{s}): && \text{state transition function}\nonumber\\
	T^{\xi}: && \text{delay between schedule and occurrence.}\nonumber
\end{eqnarray}
Condition to schedule and to cancel an event are logic expressions based on the system state $\textbf{s}$, and the output will be true if the system state allows to schedule or cancel executions. The occurrence of an event changes the system state, which is indicated by the state transition function. The input and output of $F^{\xi}(\textbf{s})$ are the system state before and after event occurrence, respectively. There is a delay, zero or positive, between scheduling time and occurring time, and this delay is called \textit{occurrence delay} and denoted as $T^{\xi}$ in the rest of the paper. In this work, $T^{\xi}$ is considered as random variate. Generally, each event usually occurs more than once during a simulation run, and the term \textit{execution} is introduced to specify a certain occurrence of an event. An execution is composed by a set of four objects, $(\xi,i,\tau_0,\tau_1)$, where $\xi$ denotes the event type, $i$ denotes the index of execution, $\tau_0$ denotes the scheduling time and $\tau_1$ denotes the occurring time.

Given a set of events $\mathbb{X}$, initial system state $\textbf{s}_0$, and simulation length defined as the number of iterations and denoted by $K$, a DES model could be implemented with the event-scheduling logic, as in Algorithm 2. The algorithm is initialized with simulation the clock equal to zero, number of executions scheduled in the past for each event type equal to zero, and empty future event list. The \textit{future event list} is a list of executions, which are scheduled but not occur yet. Each iteration is composed by three steps, which are scheduling new executions, canceling executions and performing an execution. First, for each event type $\xi$, if the condition to schedule $\xi$ is true, a new execution of $\xi$ is created and added to the future event list, as in lines 3 to 9. Then, for each execution in the future event list, if the condition to cancel the event is true, the executions of that event type is removed from the future event list, as in lines 12 to 16. Finally, the execution with the earliest occurring time is taken from the future event list, the system state is changed according to the state transition function of the event type, and the simulation clock is set to the occurring time. If there are more than one execution with the same earliest occurring time, the algorithm can randomly pick one among them or according to a priority rule.
%The event-scheduling approach is the logic behind all the major DES software and used by practitioners when developing simulation codes with general purpose languages \citep{law2014simulation}. For sake of completeness, the logic is briefly described as shown in Figure \ref{Fig:SimAlgo}. The fundamental elements are the system state and the events. The system state is a collection of state variables to describe the system at a particular time, while an event is everything whose execution can change the system state. All the already scheduled but not yet executed events are stored in the future event list together with their occurring times. When simulation is launched, the system state and the future event list is initialized with user-defined values, and the simulation clock is set to zero. The event with the earliest occurring time in the future event list will be executed, and the system evolves into a new state together with the simulation clock. The new state may enable to schedule new events, i.e., adding new event together with execution time to the future event list, or cancel events, i.e., removing some event executions from the future event list. There is usually a delay between the time when an event is added to the event list and the time when the event is executed. In the following, we refer the time when an event is added to the event list as the \textit{scheduling time}, and the time when an event is executed as the \textit{execution time}. The algorithm will terminate when some given condition is met or when a given value for the simulation clock is reached.

%\begin{figure}[h]
%	\centering
%	\includegraphics[width=0.6\textwidth]{Figures/EventSimAlgo.png}
%	\caption{Event-scheduling simulation algorithm.}
%	\label{Fig:SimAlgo}
%\end{figure}

%\section{Mathematical programming representation of DES}
In the following, a procedure to translate DES models into MPR models, based on the event-scheduling logic, is introduced. Before presenting the procedure, the assumptions that the DES model has to satisfy in order to have the procedure applicable, are introduced.

\subsection{Assumptions}
To apply the procedure proposed in this work, the following assumptions must be satisfied.

\begin{enumerate}
	\item State variables are integer.
	\item For all the event types $\xi$, the \textit{condition to schedule} and the \textit{condition to cancel} are in the form $a^{\xi,s}\le s \le c^{\xi,s}$, where $a^{\xi,s}$ and $c^{\xi,s}$ are integers and represent, respectively, the lower bound and the upper bound of a range, for a certain state variable $s$. When multiple state variables are involved, they are combined using the logical operator ``AND".
	%and combined with logic operator ``AND" when multiple state variables are involved, where $a^{\xi,s}$ and $b^{\xi,s}$ are given.
	%\item The scheduling condition of an event is independent of the history and not changed along time.% (It could be possible to define more state variables in case of history dependence and time-variant scheduling conditions.)
	%\item An event execution of $e^{\xi}$ leads to (integer) increment or decrement equal to $\Delta^{\xi}_s$ of certain state variables $s$, and $\Delta^{\xi}_s$ is not changed along time. (A direct evaluation can be modeled in this way.)
	\item The occurring delay of each execution can be generated off-line. %(\textit{This point is different from ERG. In ERG, the delay is dependent on the edge, i.e, a couple of events, but I consider delay dependent on a single event.})
	\item When more than one execution in the future event list have the same occurring time% as the earliest execution time
	, the sequence of performing the executions is immaterial, i.e., different sequences lead to the same new system state and the same new event list when the simulation clock is advanced.
	\item Only events with strictly positive occurrence delay can be canceled.
	%	the system state, the event list and the number of past executions of each event type are independent of the sequence of execution when the simulation clock is advanced.
	%\item Simulation terminates when the total number of executions reaches a known value $K$.
	%Simulation terminates when the number of execution of each event $e^{\xi}$ reaches $N^{\xi}$, which is known, i.e., the number of entities in the system is known. 
	%\item Only event with positive-delay between scheduling and execution can be canceled.
\end{enumerate}

The first assumption requires the state variables to be integer. Integer variables widely exist in DES models, such as number of jobs in buffers, idle servers, and binary variables to model system behavior and control. A discrete state variable can be translated into an integer variable or a set of binary variables. For real-valued state variables, they can be approximately discretized. Thus, the first assumption is fairly general.

The second assumption is, instead, more strict. However, if a model does not satisfy this condition, one can consider to introduce extra binary variables to satisfy assumption (2). For instance, if the condition to schedule executions of event type ${\xi}$ is $s\le a^{\xi,s} \ OR\ s\ge c^{\xi,s}$, a binary variable $s^{'}$ can be introduced in the model, and $s^{'}$ is equal to one if and only if 
$s\le a^{\xi,s} \ OR\ s\ge c^{\xi,s}$. Thus, violation of assumption (2) can be overcome, even though at the cost of an increase of the model complexity.

As for the third assumption, the delay represents usually inter-arrival time or service time in queueing systems. These delays will be used as the constant term of some constraints in the proposed MPR, hence, they should be generated offline, i.e., independently from the real-time state of the system. For instance, both stationary and non-stationary arrival process will satisfy this assumption.
%For example, if an event represents the finishing of a job on a machine, it is scheduled (i.e., put in the list) when the job starts its processing, and it will be executed when the jobs will be finished, i.e., the time between scheduling and execution is the job processing time. 
%If the condition is not met, i.e., if the occurrence delay is not iid, it can be splitted into several events so that each event has iid delay. For instance, if the distribution of service time depends on the job type, the event of \textit{finishing} a job should be splitted such that the \textit{finishing} of each job type is represented by one event.

The fourth assumption, in practice, says that the execution time is the only attribute of priority for the event executions in the future event list. If one would like to specify a different priority for events having the same execution time, he/she can specify the priority by adding the event with lower priority to the event list after the execution of the event with higher priority, which can be done by introducing extra binary variables.

The fifth assumption is an extension of the fourth one. Assumption (4) implies that only events with positive delay can be canceled. Let us assume that there is a zero-delay event ${\xi}$ with cancellation. After an execution, $(\xi,i,\tau_0,\tau_0,false)$, is added to the future event list in iteration $k$. It is obvious that the condition to schedule and the condition to cancel an event are not identical; thus, either the new execution must occur immediately if there is no other execution in the future event list having the same occurring time $\tau_0$ and event cancellation is not relevant, or there are such executions and two cases can occur. In one case, execution $(\xi,i,\tau_0,\tau_0,false)$ is immediately executed in iteration $(k+1)$. In the other case, if some other executions with occurring time equal to $\tau_0$ are executed earlier than execution $(\xi,i,\tau_0,\tau_0,false)$, and the condition to cancel $\xi$ is true at a certain time, then the execution $(\xi,i,\tau_0,\tau_0,false)$ is canceled and never occurs. The simulation realization in the latter case is different from the former one, and assumption (4) is violated. Thus, only positive-delay event can be canceled.
%The fifth condition specifies termination condition of the event-scheduling algorithm. %Specifically, for queueing systems, the termination condition usually refers to the number of jobs passing through each station.

%The proposed procedure cannot handling event cancellation, as stated in the last assumption.

\subsection{Modified event-scheduling execution logic} \label{sec:MPR_procedure}

It is not trivial to generate an MPR based on Algorithm 2. Thus, a modified event-scheduling execution logic is proposed this modified algorithm is equivalent to Algorithm 2, but enables the MPR generation. In the modified algorithm, each event is categorized as zero-delay or positive-delay, depending on whether the occurrence delay $T^{\xi}$ is equal to zero with probability equal to one or not. The modifications differ for zero-delay and positive-delay events.

In Algorithm 2, there may be more than one execution of the event type $\xi$ in the future event list. The first modification is that the number of executions of any zero-delay event in the future event list at any time is no more than one. Multiple executions are equivalent to sequentially schedule a new execution after the previous one occurs.

The second modification is about the scheduling of positive-delay events. A \textit{counting event} ${\tilde{\xi}}$ and a counting variable $u^{\xi}$ are artificially created. The counting event ${\tilde{\xi}}$ is zero-delay, with the same scheduling condition as ${\xi}$. The simulation logic is then modified as follows. When the system state satisfies the condition to schedule $\xi$, an execution of event $\tilde{\xi}$, other than $\xi$, is added to the future event list. Only after event ${\tilde{\xi}}$ occurs, an execution of event type ${\xi}$ is added to the future event list. The counting variable $u^{\xi}$ represents the number of executions of ${\xi}$ in the future event list, and its value is incremented by one when counting event ${\tilde{\xi}}$ occurs and decremented by one when event ${\xi}$ occurs.

The third modification is about event cancellation. A fifth object named \textit{cancel} is introduced to an \textit{execution}, which has a binary value indicating if the execution is canceled. It is initialized as \textit{false} (i.e., zero) when an execution is created. Once the condition to cancel executions of event type $\xi$ is true, the value of \textit{cancel} of those executions is changed to \textit{true}, (i.e., one), the counting variable $u^{\xi}$ is set to zero, but the execution is not removed from the future event list. When a canceled execution is taken from the future event list, the system state will remain unchanged and the algorithm will continue to the next iteration.

Based on the assumptions and the three modifications above, compositions of zero-delay and positive-delay events can be modified accordingly. Zero-delay events are composed by a collection of two objects ($CS^{\xi}(\mathbf{s})$, $F^{\xi}(\mathbf{s})$), which are condition to schedule and state transition. Positive-delay events are composed by five objects ($T^{\xi}$, $\tilde{\xi}$, $u^{\xi}$, $CC^{\xi}(\mathbf{s})$, $F^{\xi}(\mathbf{s})$), which are occurrence delay, counting event, counting variable, condition to cancel and state transition, respectively.

The last modification is that besides integer $K$, which represents the total number of iterations, the number of executions of each event $\xi$, denoted by $N^{\xi}$, should also be provided. It is not necessary that $N^{\xi}$ is exactly equal to the number of executions of event $\xi$ in the simulation run, instead, it could be an upper bound. Generally speaking, $N^{\xi}$ can be equal to K, but the smaller the value, the smaller the number of variables, thus, a simpler model will be consequently developed. Since $N^{\xi}$ is not the exact number of executions of event $\xi$, when the simulation terminates, the future event list sometimes can be not empty, i.e., some executions are scheduled but never occur.

The modified event-scheduling logic is shown in Algorithm 3. The differences between the two algorithms are the following. In each iteration, only zero-delay events are scheduled by validating the condition to schedule, as in line 3. Positive-delay events are scheduled after the execution of its counting event is performed, as in lines 26 to 30. Canceled executions are not removed from the future event list, but the attribute \textit{cancel} is set to true, and the counting variable is set zero, as in line 13. When performing an execution, only the transition function of uncanceled executions will be applied, as in lines 21 to 23.

%The condition to schedule, the condition to cancel, the distribution of delay between scheduling and execution $T^{\xi}$ should satisfy the assumptions (2) and (3). In the future event list, there is usually an upper bound on number of executions of the each event $e^{\xi}$, represented by $\beta^{\xi}$. For instance, to simulate an arrival process, the maximum number of arrivals in the event list is equal to $1$, since a new arrival can only be scheduled by the previous one, with a delay equal to the inter-arrival time. In the future event list, there may be one or more executions of the same event $e^{\xi}$. Events with $\beta^{\xi}$ equal to one, are named \textit{single-execution} events, and events with $\beta^{\xi}$ greater than one are named \textit{multi-execution} events. All the zero-delay events are defined as single-execution because, when the delay between scheduling and execution of an event is zero, multiple executions are equivalent to the sequential scheduling of a single-execution event. For positive-delay events $e^{\xi}$, the following procedure is introduced. A \textit{counting} event ${e}^{\tilde{\xi}}$ and counting variable $u^{\xi}$ are created artificially. The counting event ${e}^{\tilde{\xi}}$ is zero-delay, with the same scheduling condition as $e^{\xi}$. Each time the condition to schedule is true, an execution of ${e}^{\tilde{\xi}}$ is added to the future event list. When ${e}^{\tilde{\xi}}$ is actually executed, an execution of $e^{\xi}$ is added to the future event list. Then, if the condition to schedule event $e^{\xi}$ is still true, another ${e}^{\tilde{\xi}}$ is scheduled, therefore, simultaneous scheduling of multiple executions of event $e^{\xi}$ can be done in a sequential way without advancing simulation clock. The counting variable $u^{\xi}$ represents the number of executions of $e^{\xi}$ in the future event list, and its value is incremented by one when counting event $e^{\tilde{\xi}}$ is executed and decremented by one when event $e^{\xi}$ is executed. If event $e^{\xi}$ is canceled, $u^{\xi}$ is reset to zero. Since the maximum number of executions of $e^{\xi}$ is equal to $\beta^{\xi}$, the inequality $u^{\xi}\le\beta^{\xi}-1$ should be included into condition to schedule it, as well as event $e^{\tilde{\xi}}$. Besides integer $K$, which represents the total number of executions before simulation termination, the number of executions of each event $\xi$ should also be provided, denoted by $N^{\xi}$. It is not necessary that $N^{\xi}$ is exactly equal to the number of executions of event $\xi$ in the simulation run, instead, it could be an upper bound of it. Generally speaking, $N^{\xi}$ can be equal to K, but the smaller the value, the smaleler the number of variables that are, thus a simpler model will be developed consequently.

To summarize, to implement the modified event-scheduling logic, zero-delay events, including counting events of positive-delay events, must be provided together with the condition to schedule and the transition function, and positive-delay events must be provided together with occurrence delay, counting event, counting variable, condition to cancel and state transition function. An example of G/G/m queue is shown in Table \ref{tab:ggm_1}. Three events and two state variables are essential. The three events represent job arrival, service start and service finish, and are denoted by ${arr}$, ${ss}$ and ${sf}$, respectively. The two state variables represent the number of jobs in the queue and the number of occupied servers, and are denoted by $q$ and $g$, respectively. State variable $q$ is non-negative integer, and $g$ can be any integer between $0$ and $m$, which is the number of servers. Since event cancellation is not relevant to G/G/m queue, the condition to cancel is not showed. Event ${arr}$ is a positive-delay event, as explained above. %and one execution can be scheduled as soon as the previous job arrives with a delay equal to a sample from inter-arrival time distribution $T^{arr}$. 
A counting event of ${arr}$, i.e., ${\tilde{arr}}$, and a counting variable $u^{arr}$ are defined. The condition to schedule ${arr}$ is that $u^{arr}$ is equal to zero, since it is a renewal process, i.e., each time a job arrives, one and only one new arrival can be scheduled. When ${arr}$ is executed, one job arrives to the system, and queue level $q$ is increased by one. Event ${ss}$ is a zero-delay event. The condition to schedule ${ss}$ is that there is one job waiting in the queue and one server available. Upon execution of ${ss}$, queue level is reduced by one, and the number of occupied servers is increased by one. Event ${fs}$ is a positive-delay event, it is scheduled immediately after ${ss}$ is executed, and the delay until its execution is equal to the service time, i.e., sampled from $T^{sf}$. Therefore, event ${ss}$ can be regarded as the counting event of ${sf}$. Furthermore, state variable $g$ is the counting variable of ${sf}$. %If the simulation terminates with $N$ jobs processed by the system, the total number of executions of all events is equal to $N$.
The only information missing in the Table to complete the algorithm is number of iterations $K$ and number of executions of each event type $N^{\xi}$.

%To derive the MPR of a simulation run of any system, one just needs state variables, zero-delay events and positive delay events as in Table \ref{tab:ggm_1}, number of iterations $K$ and number of executions $N^{\xi}$. Then, the MPR is derived as in next sections.

%To implement the event-schedule algorithm and generate MPR, events and state variables should be first defined. 
%For instance, to simulate a G/G/m system, three events and two state variables are essential. The three events represent job arrival, service start and service finish, and are denoted by ${arr}$, ${ss}$ and ${sf}$, respectively. The two state variables represent the number of jobs in the queue and the number of occupied servers, and are denoted by $q$ and $g$, respectively. State variable $q$ is non-negative integer, and $g$ can be any integer between $0$ and $m$, which is the number of servers.

%To implement the event-scheduling logic in Algorithm 1, for each event $\xi$, the condition to schedule $CS^{\xi}$, the condition to cancel $CC^{\xi}$, the distribution of occurrence delay $T^{\xi}$ %, the maximum number of executions in the future event list $\beta^{\xi}$ and the state variable changes upon execution are also necessary. 
%and the state transition function must be provided. Besides, to derive the MPR with the proposed approach, some modifications must be made to the general event-scheduling algorithm in Algorithm 1, but the resulting algorithm will be equivalent. Each event is categorized as zero-delay or positive-delay, depending on whether the occurrence delay $T^{\xi}$ is equal to zero with probability equal to one or not. The modifications differ for zero-delay and positive-delay events.

>>>>

\bibliographystyle{apacite}
\bibliography{RAP, Biblio_Loop}
